---
title: "Data Cleaning"
output: github_document
---


## Session Setup

```{r setup}
knitr::opts_chunk$set(echo = TRUE)

## Packages
library(tidyverse)     # Tools and syntax
library(peacesciencer) # Conflict and dyadic data
library(WDI)           # World Bank data
library(countrycode)   # To help cross-walk datasets
```


## Reading in Data

### ODA/Foreign Aid

First, let's deal with ODA data. As noted below, this comes from the OECD.stat database and consists of total ODA commitments from DAC member countries to developing countries. It covers 1966 to 2021.

```{r}
## OECD Aid Data
## - Downloaded 05-03-2023
## - Aid (ODA) commitments to countries and regions [DAC3a]
## - Total commitments in 2021 US dollars
oecd <- read_csv(
  here::here(
    "01_data",  
    "oecd_oda_commitments.csv"
  )
)
```

With the data read in, I next need to do some cleaning to help cross-walk it with other datasets.

```{r}
glimpse(oecd)
oecd %>%
  transmute(
    donor = countrycode(
      Donor,
      "country.name",
      "country.name"
    ),
    recipient = countrycode(
      Recipient,
      "country.name",
      "country.name"
    ),
    year = Year,
    oda_millions = Value
  ) %>%
  drop_na(recipient, donor) -> oecd
summary(oecd)
```

Alright, there are probably some missing dyads here. Let's fix that by making a set of all possible combinations of donors and recipients over time:

```{r}
dry <- expand_grid(
  donor = unique(oecd$donor),
  recipient = unique(oecd$recipient),
  year = unique(oecd$year)
)
dim(dry) # should by 299,040 dyad-years
```

Okay, let's join together:

```{r}
oecd <- left_join(dry, oecd)
```

One final thing. Let's sort this by years.

```{r}
oecd <- arrange(oecd, year)
```

Oh, and I should probably make an indicator to alert me to when certain donors are inactive. I'd also like to convert NAs to zero.

```{r}
oecd %>%
  group_by(donor, year) %>%
  mutate(
    donor_active = ifelse(
      sum(!is.na(oda_millions)) > 0,
      1, 0
    )
  ) %>%
  ungroup() %>%
  mutate(
    oda_millions = replace_na(oda_millions, 0)
  ) -> oecd
```

Okay that's done. Let's move to some dyadic data.

### Conflict and dyadic data

I'm using data accessible via the `{peacesciencer}` package to get data on conflict and other dyadic factors relevant for the analysis.

Here's the code to pull together the relevant data:

```{r}
## Create dyad-year base (this is directed by default)
create_dyadyears(
  subset_years = 1966:2021
) %>%
  
  ## include GW codes, too
  add_gwcode_to_cow() %>%
  
  ## Adding in some covariates
  add_atop_alliance() %>%
  add_capital_distance() %>%
  add_cow_trade() %>%
  add_democracy() %>%
  add_sdp_gdp() %>%
  add_fpsim() %>%
  add_igos() -> ps_data
```

Annoyingly I have to do conflict data at the state-year level. Below I pull together the relevant data. Then I'll merge to the data above.

```{r}
create_stateyears(
  subset_years = 1996:2021
) %>%
  add_gwcode_to_cow() %>%
  add_ucdp_acd(
    type = c("intrastate",
             "II")
  ) -> ucdp_data
```

Let's prep for some cross-walking:

```{r}
ps_data %>%
  mutate(
    donor = countrycode(
      ccode1,
      "cown",
      "country.name"
    ),
    recipient = countrycode(
      ccode2,
      "cown",
      "country.name"
    )
  ) -> ps_data
ucdp_data %>%
  mutate(
    recipient = countrycode(
      ccode,
      "cown",
      "country.name"
    )
  ) -> ucdp_data
```

Let's merge these:

```{r}
ps_data <- left_join(ps_data, ucdp_data)
```

### Some gravity/distance data

I want some covariates for colonial past and maybe an alternative measure of distance. Here's come code to download and access data from CEPII.

```{r}
distance_geo <- haven::read_dta(
  here::here("01_data","dist_cepii.dta")
)
distance_geo %>%
  mutate(
    donor = countrycode(
      iso_o,
      "iso3c",
      "country.name"
    ),
    recipient = countrycode(
      iso_d,
      "iso3c",
      "country.name"
    )
  ) -> distance_geo
```


Alright, on to the World Bank!


### World Bank data

What I really need here is a measure related to disasters. Thankfully, the WB has a variable on the number of displaced persons due to natural disasters.

```{r}
# WDIsearch("disaster") %>% as_tibble() %>% view
# I need: VC.IDP.NWDS
WDI(
  country = "all",
  indicator = "VC.IDP.NWDS",
  start = 1966
) -> disaster_data

## Some recodes
disaster_data %>%
  transmute(
    recipient = countrycode(
      country,
      "country.name",
      "country.name"
    ),
    year,
    displaced_disaster = VC.IDP.NWDS
  ) %>%
  drop_na(recipient) -> disaster_data

## convert NAs to zero

disaster_data %>%
  mutate(
    displaced_disaster = replace_na(
      displaced_disaster, 0
    )
  ) -> disaster_data
```

### Some migration data

```{r}
mig_data <- read_csv(
  here::here("01_data", "oecd_migration_stock.csv")
)
mig_data %>%
  transmute(
    donor = countrycode(
      Country,
      "country.name",
      "country.name"
    ),
    recipient = countrycode(
      `Country of birth/nationality`,
      "country.name",
      "country.name"
    ),
    year = Year,
    migrants = Value
  ) -> mig_data
mig_data <- mig_data %>%
  drop_na(donor, recipient)
```


## Combining

Okay, let's work on combining these.

```{r}
## Merge peacesciencer data with aid data
full_dt <- left_join(oecd, ps_data)
## check dimensions
nrow(full_dt) == nrow(oecd) # good

## Merge with gravity/distance data
full_dt <- left_join(full_dt, distance_geo)
## check dimensions
nrow(full_dt) == nrow(oecd) # good

## Merge with disaster data
full_dt <- left_join(full_dt, disaster_data)
## check dimensions
nrow(full_dt) == nrow(oecd) # good

## Merge with migration data
full_dt <- left_join(full_dt, mig_data)
## check dimensions
nrow(full_dt) == nrow(oecd)
```

## Final cleaning

Alright, let's put the final coat of paint on this. Basically, I want to just clean up some variable names and deal with some possible missing values.

```{r}
## Drop some redundant country codes
full_dt %>%
  select(
    -c(ccode1:gwcode2, ccode:gwcode, conflict_ids:iso_d)
  ) -> full_dt

## Oh, and drop some values for donors
full_dt %>%
  mutate(
    imports = flow1,
    exports = flow2
  ) -> full_dt
```

Time for some imputation using random forests. To see more about the method look at the help file for missRanger(). Be patient with this bit.

```{r}
# library(missRanger)
# imputed_dt <- missRanger(
#   full_dt,
#   pmm.k = 3
# )
```


## Save

I'll save the final data in the 01_data folder:

```{r}
write_csv(
  full_dt,
  here::here("01_data", "final_data.csv")
)
```

